Estimating nucleon substructure properties in a\\[.2ex] unified model of p-Pb and Pb-Pb collisions

### 2018-05-10 talk/qm2018.tex
update figures

\tikzset{
  box/.style={
    align=center,
    inner sep=1ex,
    background default fill=black!10,
    background default text=black!30,
    background fill=theme!10,
    background text=theme,
  }
}

@@

\begin{frame}[t]{\scshape Bayesian parameter estimation}
  \makebox[\textwidth]{
    \begin{tikzpicture}[overlay]
      \node[box=1, font=\small] (input) at (-5, -1) {
      \boxtitle{Input parameters}
      IC and QGP properties
    };
    \node[box, below=of input.south west, anchor=north west, yshift=.6cm, font=\small] (physics) {
      \boxtitle{Physics theory}
      initial stages, hydro, and\\
      Boltzmann transport
    };
    \node[box, right=of input.east, anchor=west, xshift=-.2cm, font=\small] (model) {
      \boxtitle{Computer model}
      minimum bias event-by-\\
      event simulations
    };
    \node[box, right=of model.east, anchor=west, xshift=-.2cm, font=\small] (gp) {
      \boxtitle{Gaussian process emulator}
      surrogate model
    };
    \node[box, below=of gp.south, anchor=north, font=\small] (mcmc) {
      \boxtitle{MCMC}
      calibrate model to data
    };
    \node[box, below=of mcmc.south, anchor=north, font=\small] (posterior) {
      \boxtitle{Posterior distribution}
      quantitative estimates \\
      of each parameter
    };
    \node[box, left=of mcmc.west, anchor=east, font=\small] (exp) {
      \boxtitle{Experimental data}
      yields, mean $p_T$, flows
    };
    \begin{scope}[color=black!70, ->, thick]
      \draw (input) -- (model);
      \draw ([yshift=-.2cm]physics.north east) -| (model.south);
      \draw (model) -- (gp);
      \draw (gp) -- (mcmc);
      \draw (exp) |- (mcmc);
      \draw (mcmc) -- (posterior);
    \end{scope}
  \end{tikzpicture}
  }
  \begin{textblock}{10}(1, 10)
    {\scshape Objective:}
    \begin{itemize}
        \item Explore parameter space of a given model
        \item Quantify posterior probability of every parameter region
          \textit{given the model, data and \textbf{known uncertainties}}
      \end{itemize}
  \end{textblock}
\end{frame}

\begin{frame}[t]{\scshape Heavy-ion collision model}
  \centering
  \medskip
  \begin{tikzpicture}
    \node (a) {\includegraphics[width=.95\textwidth]{hannah_collision}};
    \node[align=left, anchor=north west] at (a.north west) {%
      \scriptsize Figure: H. Petersen, MADAI
      };
  \end{tikzpicture}
  \begin{columns}[T]
    \begin{column}{0.3\textwidth}
      \texttt{\scshape\color{theme} TrENTo} {\small initial conditions}\\
      \smallskip\hrule\medskip
      {\scriptsize parametric entropy deposition}\\
      {\tiny \color{theme}PRC.92.011901}\\[2ex]
      \texttt{\scshape\color{theme} freestream} {\small pre-flow}\\
      \smallskip\hrule\medskip
      {\scriptsize infinitely weak coupling limit}\\
      {\tiny \color{theme} PRC.91.064906, ~PRC.80.034902}
    \end{column}
    \hfill
    \begin{column}{0.3\textwidth}
      \texttt{\scshape\color{theme} VISH2+1} {\small viscous hydro}\\
      \smallskip\hrule\medskip
      {\scriptsize 14-mom.\ approx w/ shear \& bulk}\\
      {\tiny \color{theme} PRC.77.064901, ~J.CPC.2015.08.039}\\[2ex]
      \texttt{\scshape\color{theme}frzout} {\small sampler}\\
      \smallskip\hrule\medskip
      {\scriptsize non-RTA shear \& bulk corrections}\\
      {\tiny \color{theme} (Bernhard thesis) arXiv:1804.06469}
    \end{column}
    \hfill
    \begin{column}{0.3\textwidth}
      \texttt{\scshape\color{theme} UrQMD} {\small Boltzmann cascade}
      \smallskip\hrule\medskip
      {\scriptsize hadronic afterburner, simulate scatterings and decays}\\
      {\tiny\color{theme} PPNP.98.00058, ~JPG.25.9.308}\\[2ex]
      {\scriptsize Final observables calculated as similar as possible to experiment}
    \end{column}
  \end{columns}
  \begin{tikzpicture}[overlay]
    \draw (-3, -.2) -- (-2.7, -.2);
    \draw (-2.7, -.2) -- (-2.7, 3.6);
    \draw (-2.7, 3.6) -- (-2.4, 3.6);
    \draw (2.4, -.2) -- (2.7, -.2);
    \draw (2.7, -.2) -- (2.7, 3.6);
    \draw (2.7, 3.6) -- (3.0, 3.6);
  \end{tikzpicture}
\end{frame}

@@

\begin{frame}[t]{\scshape Participant thickness functions}
  \begin{columns}[T]
    \begin{column}{.5\textwidth}
      \begin{figure}
        \includegraphics[width=\textwidth]{trento_thickness}\\
        \small Participant matter
      \end{figure}
    \end{column}
    \begin{column}{.5\textwidth}
      \centering \bigskip
      \textcolor{theme}{Original \trento\ model:}
      \begin{equation*}
        T_\text{nucleus}(x, y) = \sum\limits_{i=1}^{N_\text{part}}
        \gamma_i\, T_\text{proton}(x - x_i, y - y_i)
      \end{equation*}
      {\small Random weight $\gamma_i$ sampled from Gamma distribution with unit mean
      and variance $1/k$.}\\[2ex]
      \textcolor{theme}{This work:}
      \begin{equation*}
        \gamma_i\, T_\text{proton} \rightarrow \sum\limits_{j=1}^{N_\text{parton}}
        \gamma_j\,T_\text{parton}
      \end{equation*}
    \end{column}
  \end{columns}
\end{frame}

@@

\begin{frame}[t]{\scshape Computer experiment design}
  \bigskip
  \begin{columns}[T]
    \begin{column}{.6\textwidth}
      \scriptsize
      \begin{tabular}{lll}
        Parameter         & Description                        & Range           \\
        \paddedhline
        Norm              & Overall normalization              & 12--28          \\
        $p$               & Entropy deposition parameter       & $-1$ to $+1$    \\
        $\sigmaf$         & Relative parton fluct. std.\       & 0--2            \\
        $w$               & Parton sampling radius             & 0.4--1.2 fm     \\
        $\X$              & Parton structure parameter         & 0--1            \\
        $\npartons$       & Number of partons                  & 1--10           \\
        $\dmin$           & Nucleon exclusion volume           & 0--4.9 fm$^3$   \\
        $\tau_{fs}$       & Free streaming time                & 0.1--1.5 fm/$c$ \\
        $\eta/s$ min      & Shear viscosity at $T_c$           & 0--0.2          \\
        $\eta/s$ slope    & Slope above $T_c$                  & 0--8 GeV$^{-1}$ \\
        $\eta/s$ crv      & Curvature above $T_c$              & $-1$ to $1$     \\
        $\zeta/s$ norm    & Bulk viscosity peak height         & 0--0.1          \\
        $\zeta/s$ width   & Bulk viscosity peak width          & 0--0.1 GeV      \\
        $\zeta/s$ temp    & Bulk viscosity peak location       & 150--200 MeV    \\
        $T_\text{switch}$ & Particlization temperature         & 135--165 MeV    \\
      \end{tabular}
    \end{column}
    \begin{column}{.4\textwidth}
      \textcolor{theme}{\scshape Important details:}
      \smallskip
      \begin{itemize}
        \small
        \item Recast constituent width $w$ as:\\[1ex]
          $w = w_\text{min} + \X(r - w_\text{min})$\\[1ex]
          $v_\text{min} = 0.2$~fm,~ $v_\text{max} = w$\\[1ex]
        \item Parton sampling radius is \emph{not} equivalent to the proton radius
          in the proton c.o.m.\ frame, see\\
          \textcolor{theme}{\scriptsize PRC.94.024919}\\[1ex]
        \item Other parameters same as combined Pb+Pb analysis at 2.76 and 5.02 TeV
      \end{itemize}
    \end{column}
  \end{columns}
  \begin{center}
    \color{theme}
    500 design points per collision system, $\mathcal{O}(10^4)$ events per design point!
  \end{center}
\end{frame}

\begin{frame}[plain]
  \begin{tikzpicture}[remember picture, overlay]
    \node[anchor=west] at (current page.west) {
      \includegraphics[height=.95\paperheight]{observables_design}
      };
    \node[anchor=east] at (current page.east) {
      \includegraphics[height=.95\paperheight]{observables_posterior}
      };
    \draw[<-] (current page.center) ++(-.75, 1.7) --
      node[above] {\small Prior}
      node[below] {\scriptsize (training data)}
      ++(1.5, 0);
    \draw[->] (current page.center) ++(-.75, -1.2) --
      node[above] {\small Posterior}
      node[below] {\scriptsize (samples)}
      ++(1.5, 0);
  \end{tikzpicture}
\end{frame}

@@

\begin{frame}[plain]
  \begin{columns}
    \begin{column}{.5\textwidth}
      Right: maximum a posteriori (best fit) predictions from the calibrated emulator; \emph{not a model calculation}.\\[2ex]
    \end{column}
    \begin{column}{.4\textwidth}
      \includegraphics[height=.95\textheight]{find_map}
    \end{column}
  \end{columns}
\end{frame}

>>>>

\tikzset{
  box/.style={
    align=center,
    inner sep=1ex,
    background default fill=black!10,
    background default text=black!30,
    background fill=theme!20,
    background text=offblack,
  }
}

@@

\begin{frame}[t]{\scshape Bayesian parameter estimation}
  \makebox[\textwidth]{
    \begin{tikzpicture}[overlay]
      \node[box=1, font=\small] (input) at (-5, -1) {
          \boxtitle{\color{theme} Input parameters}
      IC and QGP properties
    };
    \node[box, below=of input.south west, anchor=north west, yshift=.6cm, font=\small] (physics) {
        \boxtitle{\color{theme} Physics theory}
      initial stages, hydro, and\\
      Boltzmann transport
    };
    \node[box, right=of input.east, anchor=west, xshift=-.2cm, font=\small] (model) {
        \boxtitle{\color{theme} Computer model}
      minimum bias event-by-\\
      event simulations
    };
    \node[box, right=of model.east, anchor=west, xshift=-.2cm, font=\small] (gp) {
        \boxtitle{\color{theme} Gaussian process emulator}
      surrogate model
    };
    \node[box, below=of gp.south, anchor=north, font=\small] (mcmc) {
        \boxtitle{\color{theme} MCMC}
      calibrate model to data
    };
    \node[box, below=of mcmc.south, anchor=north, font=\small] (posterior) {
        \boxtitle{\color{theme} Posterior distribution}
      quantitative estimates \\
      of each parameter
    };
    \node[box, left=of mcmc.west, anchor=east, font=\small] (exp) {
        \boxtitle{\color{theme} Experimental data}
      yields, mean $p_T$, flows
    };
    \begin{scope}[color=black!70, ->, thick]
      \draw (input) -- (model);
      \draw ([yshift=-.2cm]physics.north east) -| (model.south);
      \draw (model) -- (gp);
      \draw (gp) -- (mcmc);
      \draw (exp) |- (mcmc);
      \draw (mcmc) -- (posterior);
    \end{scope}
  \end{tikzpicture}
  }
  \begin{textblock}{10}(1, 10)
    {\scshape Objective:}
    \begin{itemize}
        \item Explore parameter space of a given model
        \item Quantify posterior probability of every parameter region
          \textit{given the model, data and \textbf{known uncertainties}}
      \end{itemize}
  \end{textblock}
\end{frame}

\begin{frame}[t]{\scshape Heavy-ion collision model}
  \centering
  \medskip
  \begin{tikzpicture}
    \node (a) {\includegraphics[width=.95\textwidth]{hannah_collision}};
    \node[align=left, anchor=north west] at (a.north west) {%
      \scriptsize Figure: H. Petersen, MADAI
      };
  \end{tikzpicture}
  \begin{columns}[T]
    \begin{column}{0.3\textwidth}
      \texttt{\scshape\color{theme} TrENTo} {\small initial conditions}\\
      \smallskip\hrule\medskip
      {\scriptsize parametric entropy deposition}\\
      {\tiny \color{theme}PRC.92.011901}\\[2ex]
      \texttt{\scshape\color{theme} freestream} {\small pre-flow}\\
      \smallskip\hrule\medskip
      {\scriptsize infinitely weak coupling limit}\\
      {\tiny \color{theme} PRC.91.064906, ~PRC.80.034902}
    \end{column}
    \hfill
    \begin{column}{0.3\textwidth}
      \texttt{\scshape\color{theme} VISH2+1} {\small viscous hydro}\\
      \smallskip\hrule\medskip
      {\scriptsize 14-mom.\ approx w/ shear \& bulk}\\
      {\tiny \color{theme} PRC.77.064901, ~J.CPC.2015.08.039}\\[2ex]
      \texttt{\scshape\color{theme}frzout} {\small sampler}\\
      \smallskip\hrule\medskip
      {\scriptsize shear \& bulk corrections}\\
      {\tiny \color{theme} (Bernhard thesis) arXiv:1804.06469}
    \end{column}
    \hfill
    \begin{column}{0.3\textwidth}
      \texttt{\scshape\color{theme} UrQMD} {\small Boltzmann cascade}
      \smallskip\hrule\medskip
      {\scriptsize hadronic afterburner, simulate scatterings and decays}\\
      {\tiny\color{theme} PPNP.98.00058, ~JPG.25.9.308}\\[2ex]
      {\scriptsize Final observables calculated as similar as possible to experiment}
    \end{column}
  \end{columns}
  \begin{tikzpicture}[overlay]
    \draw (-3, -.2) -- (-2.7, -.2);
    \draw (-2.7, -.2) -- (-2.7, 3.6);
    \draw (-2.7, 3.6) -- (-2.4, 3.6);
    \draw (2.4, -.2) -- (2.7, -.2);
    \draw (2.7, -.2) -- (2.7, 3.6);
    \draw (2.7, 3.6) -- (3.0, 3.6);
  \end{tikzpicture}
\end{frame}

@@

\begin{frame}[t]{\scshape Participant thickness functions}
  \begin{columns}[T]
    \begin{column}{.5\textwidth}
      \begin{figure}
        \includegraphics[width=\textwidth]{trento_thickness}\\
        \small Participant matter
      \end{figure}
    \end{column}
    \begin{column}{.5\textwidth}
      \centering \bigskip
      \textcolor{theme}{Original \trento\ model:}
      \begin{equation*}
        T_\text{nucleus}(x, y) = \sum\limits_{i=1}^{N_\text{part}}
        \gamma_i\, T_\text{proton}(x - x_i, y - y_i)
      \end{equation*}
      {\small Random weight $\gamma_i$ sampled from Gamma distribution with unit mean
      and variance $1/k$.}\\[2ex]
      \textcolor{theme}{This work:}
      \begin{equation*}
        \gamma_i\, T_\text{proton} \rightarrow \sum\limits_{j=1}^{M}
        \gamma_j\,T_\text{constituent}
      \end{equation*}
      \small where $M$ is the number of constituents.
    \end{column}
  \end{columns}
\end{frame}

@@

\begin{frame}[t]{\scshape Computer experiment design}
  \bigskip
  \begin{columns}[T]
    \begin{column}{.6\textwidth}
      \scriptsize
      \begin{tabular}{lll}
        Parameter         & Description                        & Range           \\
        \paddedhline
        Norm              & Overall normalization              & 12--28          \\
        $p$               & Entropy deposition parameter       & $-1$ to $+1$    \\
        $\sigmaf$         & Relative parton fluct. std.\       & 0--2            \\
        $r$               & Parton sampling radius             & 0.4--1.2 fm     \\
        $\X$              & Parton structure parameter         & 0--1            \\
        $m$               & Number of constituents             & 1--10           \\
        $\dmin$           & Nucleon exclusion volume           & 0--4.9 fm$^3$   \\
        $\tau_{fs}$       & Free streaming time                & 0.1--1.5 fm/$c$ \\
        $\eta/s$ min      & Shear viscosity at $T_c$           & 0--0.2          \\
        $\eta/s$ slope    & Slope above $T_c$                  & 0--8 GeV$^{-1}$ \\
        $\eta/s$ crv      & Curvature above $T_c$              & $-1$ to $1$     \\
        $\zeta/s$ norm    & Bulk viscosity peak height         & 0--0.1          \\
        $\zeta/s$ width   & Bulk viscosity peak width          & 0--0.1 GeV      \\
        $\zeta/s$ temp    & Bulk viscosity peak location       & 150--200 MeV    \\
        $T_\text{switch}$ & Particlization temperature         & 135--165 MeV    \\
      \end{tabular}
    \end{column}
    \begin{column}{.4\textwidth}
      \textcolor{theme}{\scshape Important details:}
      \smallskip
      \begin{itemize}
        \small
        \item Recast constituent width $w$ as:\\[1ex]
          $w = w_\text{min} + \X(w_\text{max} - w_\text{min})$\\[1ex]
          $w_\text{min} = 0.2$~fm,~ $w_\text{max} = r$\\[1ex]
        \item Parton sampling radius is \emph{not} equivalent to the proton radius
          in the proton c.o.m.\ frame, see\\
          \textcolor{theme}{\scriptsize PRC.94.024919}\\[1ex]
        \item Other parameters same as combined Pb+Pb analysis at 2.76 and 5.02 TeV
      \end{itemize}
    \end{column}
  \end{columns}
  \begin{center}
    \color{theme}
    500 design points per collision system, $\mathcal{O}(10^4)$ events per design point!
  \end{center}
\end{frame}

\begin{frame}[plain]
  \begin{tikzpicture}[remember picture, overlay]
    \node[anchor=west] at (current page.west) {
      \includegraphics[height=.95\paperheight]{observables_design}
      };
    \node[anchor=east] at (current page.east) {
      \includegraphics[height=.95\paperheight]{observables_posterior}
      };
    \draw[<-] (current page.center) ++(-.75, 1.7) --
      node[above] {\small Prior}
      node[below] {\scriptsize (training data)}
      ++(1.5, 0);
    \draw[->] (current page.center) ++(-.75, -1.2) --
      node[above] {\small Posterior}
      node[below] {\scriptsize (samples)}
      ++(1.5, 0);
    \node[inner sep=1ex, anchor=north, align=center] at (current page.north) {
      \scriptsize Data: CMS \& ALICE
    };
  \end{tikzpicture}
\end{frame}

@@

\begin{frame}[plain]
  \begin{columns}
    \begin{column}{.4\textwidth}
      \begin{tikzpicture}[remember picture, overlay]
        \node[anchor=west] at (current page.west) {\includegraphics[height=\textheight]{find_map}};
      \end{tikzpicture}
    \end{column}
    \begin{column}{.6\textwidth}
      \textit{Maximum a posteriori} (best fit) predictions from the calibrated \emph{emulator} (not the model).\\[2ex]
      \begin{table}
        \scriptsize
        \begin{tabular}{lll}
          Parameter         & Description                        & Value           \\
          \paddedhline
          Norm              & Overall normalization              & 20.3            \\
          $p$               & Entropy deposition parameter       & 0.04            \\
          $\sigmaf$         & Relative constituent fluct. std.\  & 0.92            \\
          $r$               & Constituent sampling radius        & 0.86 fm         \\
          $\X$              & Constituent structure parameter    & 0.32            \\
          $m$               & Number of constituents             & 6               \\
          $\dmin$           & Nucleon exclusion volume           & 1.1 fm$^3$      \\
          $\tau_{fs}$       & Free streaming time                & 0.42 fm/$c$     \\
          $\eta/s$ min      & Shear viscosity at $T_c$           & 0.11            \\
          $\eta/s$ slope    & Slope above $T_c$                  & 0.98 GeV$^{-1}$ \\
          $\eta/s$ crv      & Curvature above $T_c$              & -0.14           \\
          $\zeta/s$ norm    & Bulk viscosity peak height         & 0.06            \\
          $\zeta/s$ width   & Bulk viscosity peak width          & 0.017 GeV       \\
          $\zeta/s$ temp    & Bulk viscosity peak location       & 0.18 MeV        \\
          $T_\text{switch}$ & Particlization temperature         & 142 MeV         \\
        \end{tabular}
      \end{table}
    \end{column}
  \end{columns}
\end{frame}
